---
title: "Data preparation"
author: "Leonardo Concetti"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r Packages, include=FALSE, results='hide'}

library(tidyverse)

```


A function to read the results file from PCIbex 

```{r}

read.pcibex <- function(filepath, auto.colnames=TRUE, fun.col=function(col,cols){cols[cols==col]<-paste(col,"Ibex",sep=".");return(cols)}) {
  n.cols <- max(count.fields(filepath,sep=",",quote=NULL),na.rm=TRUE)
  if (auto.colnames){
    cols <- c()
    con <- file(filepath, "r")
    while ( TRUE ) {
      line <- readLines(con, n = 1, warn=FALSE)
      if ( length(line) == 0) {
        break
      }
      m <- regmatches(line,regexec("^# (\\d+)\\. (.+)\\.$",line))[[1]]
      if (length(m) == 3) {
        index <- as.numeric(m[2])
        value <- m[3]
        if (is.function(fun.col)){
          cols <- fun.col(value,cols)
        }
        cols[index] <- value
        if (index == n.cols){
          break
        }
      }
    }
    close(con)
    return(read.csv(filepath, comment.char="#", header=FALSE, col.names=cols))
  }
  else{
    return(read.csv(filepath, comment.char="#", header=FALSE, col.names=seq(1:n.cols)))
  }
}

```


Loading the results file

```{r}

dat <- read.pcibex("results.csv")

```


# Data preparation


Get rid of practice phase, instructions etc

```{r}

dat <- dat |> 
  filter(Label %in% c("experimental-experimental", "experimental-filler"))

```


Get rid of the filler trials (only do this after having checked everything is ok, including accuracy of participants etc)

```{r}

# dat <- dat |> 
  # filter(condition != "fill")

```


Get rid of the questions/answers data (not necessary, use only in certain cases to make the dataset lighter)

```{r}

# dat <- dat |> 
#  filter(Newline. == "false")

```


Remove useless columns

```{r}

dat <- dat |> 
  select(-Inner.element.number, -PennElementType, -PennElementName,
         -Latin.Square.Group, -MD5.hash.of.participant.s.IP.address,
         -Newline., -Controller.name, -EventTime, -Comments, -Label,
         -STUDY_ID, -SESSION_ID, -Order.number.of.item)

```


Remove metadata from each trial starting parameter

```{r}

dat <- dat |> 
  filter(!Parameter %in% c("_Trial_", "_Header_"))

```


Convert subject IDs to numbers starting from 1

```{r}

dat <- dat |> 
  mutate(PROLIFIC_ID = as.integer(factor(PROLIFIC_ID, levels = unique(PROLIFIC_ID))))

```


Rename the subjects based on their unique reception time (only necessary in case we don't get the subject ID otherwise).

```{r}

# dat <- dat |> 
 # mutate(subj = match(Results.reception.time, unique(Results.reception.time)))

```


Now we can get rid of the original reception time column

```{r}

dat <- dat |> 
  select(- Results.reception.time)

```


Rename columns to more manageable names 

```{r}

dat <- dat |> 
  rename(roi_value = Value, rt = Reading.time, sentence = Sentence..or.sentence.MD5.,
         roi_number = Parameter, list = group, subj = PROLIFIC_ID)

```


Separate data from sentences and question/answer

```{r}

dat <- dat |> 
  mutate(selection = case_when(
    roi_value %in% c("Si", "No") ~ roi_value,
                               FALSE ~ NA_character_)
    )


```


Paste the selected answer to each row of that trial 

```{r}

 # Define a function to replace NA values with non-NA values within each group
replace_na_with_non_na <- function(x) {
  non_na_value <- na.omit(x)[1]  # Get the first non-NA value
  replace(x, is.na(x), non_na_value)  # Replace NA values with the non-NA value
}


dat <- dat |> 
  group_by(item, subj) |> 
  mutate(selection = replace_na_with_non_na(selection)) |> 
  ungroup()

```


Change "Si" into "si" and "No" into "no" to avoid confusion with capital letters

```{r}

dat <- dat |> 
  mutate(
    roi_value = case_when(
      roi_value == "Si" ~ "si",
      roi_value == "No" ~ "no",
      TRUE ~ roi_value
    ),
    selection = case_when(
      selection == "Si" ~ "si",
      selection == "No" ~ "no",
      TRUE ~ selection
    )
  )

```



Add a column with correct/incorrect answer info

```{r}

dat <- dat |> 
mutate(correct = if_else(answer == selection, 1, 0))

```


Add a column for trial ID

```{r}

dat <- dat %>%
    mutate(trial = rep(1:(ceiling(n() / 13)), each = 13)[1:n()])

```


Arrange the columns in the desired order

```{r}

dat <- dat |> 
  select(subj, list, trial, item, condition, roi_number, roi_value, rt, sentence, 
         question, question_type, answer, selection, correct, item_type)  #remember to add here any other column that you did not remove from the data frame

```


Make sure RTs are interpreted as numbers and not characters

```{r}

class(dat$rt)

dat <- dat |> 
  mutate(rt = as.numeric(rt))

class(dat$rt)

```

Check design

```{r}

xtabs(~ subj + as.numeric(item), dat)

dat$roi_value[dat$roi_number == 10 & dat$condition != "fill"] #to check that the matrix verb is always in roi number 10

dat$roi_value[dat$roi_number == 5 & dat$condition != "fill"] #to check that the RC verb is always in roi number 5

```


# Save the clean dataset 

```{r}

write_csv(dat, r"[C:\Users\lconc\OneDrive - student.unisi.it\PhD - DISPOC\Projects\2 - Year 2\coordination_retrieval\coord_spr.csv]")

```