---
title: "Data preparation"
author: "Leonardo Concetti"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r Packages, include=FALSE, results='hide'}

library(tidyverse)

```


A function to read the results file from PCIbex 

```{r}

read.pcibex <- function(filepath, auto.colnames=TRUE, fun.col=function(col,cols){cols[cols==col]<-paste(col,"Ibex",sep=".");return(cols)}) {
  n.cols <- max(count.fields(filepath,sep=",",quote=NULL),na.rm=TRUE)
  if (auto.colnames){
    cols <- c()
    con <- file(filepath, "r")
    while ( TRUE ) {
      line <- readLines(con, n = 1, warn=FALSE)
      if ( length(line) == 0) {
        break
      }
      m <- regmatches(line,regexec("^# (\\d+)\\. (.+)\\.$",line))[[1]]
      if (length(m) == 3) {
        index <- as.numeric(m[2])
        value <- m[3]
        if (is.function(fun.col)){
          cols <- fun.col(value,cols)
        }
        cols[index] <- value
        if (index == n.cols){
          break
        }
      }
    }
    close(con)
    return(read.csv(filepath, comment.char="#", header=FALSE, col.names=cols))
  }
  else{
    return(read.csv(filepath, comment.char="#", header=FALSE, col.names=seq(1:n.cols)))
  }
}

```


Loading the results file

```{r}

dat <- read.pcibex("results.csv")

```


# Data preparation


Get rid of practice phase, instructions etc

```{r}

dat <- dat |> 
  filter(Label == "experimental-trial")

```


Get rid of the filler trials

```{r}

dat <- dat |> 
  filter(condition != "fill")

```


Get rid of the questions/answers data (not necessary, use only in certain cases to make the dataset lighter)

```{r}

dat <- dat |> 
  filter(Newline. == "false")

```


Remove useless columns

```{r}

dat <- dat |> 
  select(-Inner.element.number, -PennElementType, -PennElementName,
         -Latin.Square.Group, -MD5.hash.of.participant.s.IP.address,
         -Newline., -Controller.name, -EventTime, -Comments)

```


Rename the subjects based on their unique reception time (only necessary in case we don't get the subject ID otherwise).

```{r}

dat <- dat |> 
  mutate(subj = match(Results.reception.time, unique(Results.reception.time)))

```


Now we can get rid of the original reception time column

```{r}

dat <- dat |> 
  select(- Results.reception.time)

```


Rename columns to more managable names 

```{r}

dat <- dat |> 
  rename(roi_value = Value, rt = Reading.time, sentence = Sentence..or.sentence.MD5.,
         roi_number = Parameter, list = group)

```


Arrange the columns in the desired order

```{r}

dat <- dat |> 
  select(subj, list, item, condition, roi_number, roi_value, rt, sentence)

```


Make sure RTs are interpreted as numbers and not characters

```{r}

class(dat$rt)

dat <- dat |> 
  mutate(rt = as.numeric(rt))

class(dat$rt)

```


# Save the clean dataset 

```{r}

write_csv(dat, "C:\Users\lconc\OneDrive - student.unisi.it\PhD - DISPOC\Projects\2 - Year 2\coordination_retrieval\coord_spr.csv")

```